---
title: APPLYING AND PROMOTING OPEN SCIENCE IN ECOLOGY - SURVEYED  DRIVERS AND CHALLENGES


output: rticles::plos_article
csl: plos.csl
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      error = TRUE, 
                      eval = TRUE)
library(drake)
library(kableExtra)
# library(patchwork)
library(tidyverse)
library(ggmosaic)
```



## Supplementary

This section contains all plots and tables of the supplementary material of this manuscript.


### Meta data


```{r metadata, fig.cap="Number of respondents."}
readd(metadata_plot)
  
```


```{r stacked-bar, fig.cap="Add a clever figure caption."}
readd(stacked_barplots)
  
```


### Question 1.1

Researchers with academic affiliation have engaged more in open science-related practices compared to other researchers, and more so for early-career researchers.

```{r results-q1-1}
loadd(result_Q1.1)

result_Q1.1 %>% 
  select(-model) %>% 
  kbl(booktabs = TRUE, 
      digits = 2,
      caption = "Model stats question 1.1") %>% 
  pack_rows("global", 1, 4) %>% 
  pack_rows("final", 5, 6) %>% 
  kable_styling()
  
```



### Question 1.2


OS practices are perceived to be more important in data-, code-, methods-sharing and publishing than educational tools, depending on activities engaged in.

```{r results-q1-2}
loadd(result_Q1.2)

result_Q1.2 %>% 
  select(-aspect) %>% 
  kbl(booktabs = TRUE, 
      digits = 2,
      caption = "Model stats question 1.2") %>% 
  pack_rows("data sharing", 1, 4) %>% 
  pack_rows("code sharing", 5, 10) %>% 
  pack_rows("method", 11, 16) %>% 
  pack_rows("open publishing", 17, 21) %>% 
  kable_styling()
  
```



### Question 1.3

BUT Most people have used open data and code, but have not contributed to open data and code.

```{r results-q1-3}
loadd(result_Q1.3)

result_Q1.3 %>% 
  select(-model) %>% 
  kbl(booktabs = TRUE, 
      digits = 2,
      caption = "Model stats question 1.3") %>% 
  pack_rows("global", 1, 3) %>% 
  pack_rows("final", 4, 5) %>% 
  kable_styling()
  
```



### Question 1.4

People more likely to use OS in supervision and teaching when they use it more in their own research.

```{r results-q1-4}
loadd(result_Q1.4)

result_Q1.4 %>% 
  select(-Domain) %>% 
  kbl(booktabs = TRUE, 
      digits = 2,
      caption = "Model stats question 1.4") %>% 
  pack_rows("Teaching", 1, 4) %>% 
  pack_rows("Supervision", 5, 9) %>% 
  kable_styling()
  
```



### Question 3.2

Colloquium participants having more experience with being taught open science-related practices are more likely to engage in those practices as researchers, and teachers.

```{r results-q3-2}
loadd(result_Q3.2)

result_Q3.2 %>% 
  select(-Action) %>% 
  kbl(booktabs = TRUE, 
      digits = 2,
      caption = "Model stats question 3.2") %>% 
  pack_rows("use data", 1, 3) %>% 
  pack_rows("use code", 4, 6) %>%
  pack_rows("use open publication", 7, 8) %>% 
  pack_rows("share data", 9, 11) %>%
  pack_rows("share code", 12, 14) %>% 
  pack_rows("share open publicaiton", 15, 16) %>%
  pack_rows("use edu tools", 17, 19) %>%
  pack_rows("do open review", 20, 21) %>%
  pack_rows("do outreach", 22, 23) %>%
  kable_styling()
  
```


### Question 3.3

Colloquium participants "think" OS practices are more important in their research compared to teaching and supervision.

```{r results-q3-3}
loadd(result_Q3.3)

result_Q3.3 %>% 
  kbl(booktabs = TRUE, 
      digits = 2,
      caption = "Model stats question 3.2") %>% 
  pack_rows("global", 1, 9) %>% 
  pack_rows("final", 10, 17) %>%
  kable_styling()
  
```
