---
title: "APPLYING AND PROMOTING OPEN SCIENCE IN ECOLOGY -SURVEYED  DRIVERS AND CHALLENGES"



date: "`r format(Sys.Date(), '%d %B %Y')`"

output: 
  #pdf_document
  bookdown::pdf_book:
    dev: cairo_pdf
    toc: no
    number_sections: false
    latex_engine: lualatex
  fig_caption: yes
  extra_dependencies: ["flafter"]
  header-includes:
    - \usepackage{caption}
    - \usepackage{booktabs}
    - \usepackage{longtable}
classoption: a4paper
bibliography: Paperpile - Nov 02 BibTeX Export.bib
# csl: Rmd/elsevier-harvard_rjt.csl


---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, error = TRUE, eval = TRUE)
library(drake)
# library("cairoDevice")
library(kableExtra)
# library(patchwork)
library(tidyverse)

```

## Supplementary

This section contains all plots and tables of the supplementary material of this manuscript.




### Question 1.1

Researchers with academic affiliation have engaged more in open science-related practices compared to other researchers, and more so for early-career researchers.

```{r results-q1-1}
loadd(result_Q1.1)

result_Q1.1 %>% 
  select(-model) %>% 
  kbl(booktabs = TRUE, 
      digits = 2,
      caption = "Model stats") %>% 
  pack_rows("global", 1, 4) %>% 
  pack_rows("final", 5, 6) %>% 
  kable_styling()
  
```



### Question 1.2


OS practices are perceived to be more important in data-, code-, methods-sharing and publishing than educational tools, depending on activities engaged in.

```{r results-q1-2}
loadd(result_Q1.2)

result_Q1.2 %>% 
  select(-aspect) %>% 
  kbl(booktabs = TRUE, 
      digits = 2,
      caption = "Model stats") %>% 
  pack_rows("data sharing", 1, 4) %>% 
  pack_rows("code sharing", 5, 10) %>% 
  pack_rows("method", 11, 16) %>% 
  pack_rows("open publishing", 17, 21) %>% 
  kable_styling()
  
```



### Question 1.3

BUT Most people have used open data and code, but have not contributed to open data and code.

```{r results-q1-3}
loadd(result_Q1.3)

result_Q1.3 %>% 
  select(-model) %>% 
  kbl(booktabs = TRUE, 
      digits = 2,
      caption = "Model stats") %>% 
  pack_rows("global", 1, 3) %>% 
  pack_rows("final", 4, 5) %>% 
  kable_styling()
  
```



### Question 1.4

People more likely to use OS in supervision and teaching when they use it more in their own research.

```{r results-q1-4}
loadd(result_Q1.4)

result_Q1.4 %>% 
  select(-domain) %>% 
  kbl(booktabs = TRUE, 
      digits = 2,
      caption = "Model stats") %>% 
  pack_rows("Teaching", 1, 4) %>% 
  pack_rows("Supervision", 5, 9) %>% 
  kable_styling()
  
```
