---
title: "APPLYING AND PROMOTING OPEN SCIENCE IN ECOLOGY -SURVEYED  DRIVERS AND CHALLENGES"



date: "`r format(Sys.Date(), '%d %B %Y')`"

output: 
  #pdf_document
  bookdown::pdf_book:
    dev: cairo_pdf
    toc: no
    number_sections: false
    latex_engine: lualatex
  fig_caption: yes
  extra_dependencies: ["flafter"]
  header-includes:
    - \usepackage{caption}
    - \usepackage{booktabs}
    - \usepackage{longtable}
classoption: a4paper
bibliography: Paperpile - Nov 02 BibTeX Export.bib
# csl: Rmd/elsevier-harvard_rjt.csl


---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, error = TRUE, eval = TRUE)
library(drake)
# library("cairoDevice")
library(kableExtra)
# library(patchwork)
library(tidyverse)

```

## Supporting Information

This section contains all plots and tables of the supplementary material of this manuscript.





```{r }
loadd(program.table)

program.table %>% 
   mutate(across(where(is.character), linebreak)) %>%
    kable(booktabs = TRUE,
        #escape = FALSE,
        col.names = linebreak(c("Day", "Time", "Activity", "Description", "Collaboration with"), align = "c"),
        caption = "Program for the Living Norway 2020 Colloquium"
        ) %>%
    column_spec(column = 1, width = "20px") %>% 
    column_spec(column = 2, width = "50px") %>% 
    column_spec(column = 3, width = "200px") %>%  
    column_spec(column = 4, width = "100px") %>% 
    column_spec(column = 5, width = "50px") %>%  
  kable_styling(font_size = 8, latex_options = "HOLD_position") 
  

    

```

\newpage



```{r }
loadd(survey.table.1)

survey.table.1 %>% 
   mutate(across(where(is.character), linebreak)) %>%
    kable(booktabs = TRUE,
        #escape = FALSE,
        col.names = linebreak(c("Part", "Question", "Options", "Response"), align = "c"),
        caption = "Survey questions distributed among attendees at the Living Norway 2020 Colloquium"
        ) %>%
    column_spec(column = 1, width = "5px") %>% 
    column_spec(column = 2, width = "180px") %>% 
    column_spec(column = 3, width = "180px") %>%  
    column_spec(column = 4, width = "100px") %>% 
  kable_styling(font_size = 7, latex_options = "HOLD_position") 
    

```

\newpage


```{r }
loadd(survey.table.23)

survey.table.23 %>% 
   mutate(across(where(is.character), linebreak)) %>%
    kable(booktabs = TRUE,
        #escape = FALSE,
        col.names = linebreak(c("Part", "Question", "Options", "Response"), align = "c"),
        caption = "Survey questions distributed among attendees at the Living Norway 2020 Colloquium"
        ) %>%
    column_spec(column = 1, width = "5px") %>% 
    column_spec(column = 2, width = "180px") %>% 
    column_spec(column = 3, width = "180px") %>%  
    column_spec(column = 4, width = "100px") %>% 
  kable_styling(font_size = 7, latex_options = "HOLD_position") 
    

```

\newpage


\newpage


```{r , fig.cap="Number of respondents by affiliation."}
readd(affiliation_plot)
  
```

\newpage

```{r , fig.cap="Number of respondents by contry."}
readd(country_plot)
  
```

\newpage

```{r , fig.cap="Respondents that had used open resources in own education."}
readd(learning_use_stackplot)
  
```

\newpage

```{r , fig.cap="Respondents that had shared resources openly in own education."}
readd(learning_share_stackplot)
  
```

\newpage

```{r , fig.cap="Respondents that had learned OS principles in own education."}
readd(learning_principles_stackplot)
  
```

\newpage

```{r , fig.cap="Open engamement among respondents in own learning."}
readd(learning_other_stackplot)
  
```

\newpage



```{r }
loadd(define.table)

  define.table %>%
  mutate(across(where(is.character), linebreak)) %>%
    kable(booktabs = TRUE,
        #escape = FALSE,
        col.names = linebreak(c("Code", "Inclusion criteria and definitions", "Examples"), align = "c"),
        caption = "Codes assigned to participant responses during qualitative analysis of the “define open science” prompt. Note that many of the examples used could be, and ultimately were, assigned to two or more of the codes."
        ) %>%
    kable_styling(font_size = 8, latex_options = "HOLD_position") %>% 
    column_spec(column = 1, width = "50px") %>% 
    column_spec(column = 2, width = "200px") %>% 
    column_spec(column = 3, width = "150px") 
    

```

\newpage




```{r results-q1-1}
loadd(result_Q1.1)

result_Q1.1 %>% 
  select(-model) %>%    
  kbl(booktabs = TRUE, 
      digits = 2,
      caption = "Parameter estimates, SE and z-values for covariates in models used to test prediction 1.1") %>% 
  pack_rows("global", 1, 4) %>% 
  pack_rows("final", 5, 6) %>% 
  kable_styling(latex_options = "HOLD_position")
  
```




```{r results-q1-3}
loadd(result_Q1.3)

result_Q1.3 %>% 
  select(-model) %>% 
  kbl(booktabs = TRUE, 
      digits = 2,
      caption = "Parameter estimates, SE and z-values for covariates in models used to test prediction 1.3") %>% 
  pack_rows("global", 1, 3) %>% 
  pack_rows("final", 4, 5) %>% 
  kable_styling(latex_options = "HOLD_position")
  
```

\newpage




```{r }
loadd(hinder.table)
  hinder.table %>%
  mutate(across(where(is.character))) %>%
  mutate(across(where(is.character), linebreak)) %>%
    kbl(booktabs = TRUE,
        escape = FALSE,
        col.names = linebreak(c("Code or category", "Inclusion criteria and definitions", "Examples"), align = "c"),
        caption = "Codes assigned to participant responses during qualitative analysis of the “what hinders your engagement in OS?” prompt. Note that many of the examples used could be, and ultimately were, assigned to two or more of the codes."
        ) %>% 
    column_spec(column = 1, width = "50px") %>% 
    column_spec(column = 2, width = "200px") %>% 
    column_spec(column = 3, width = "150px") %>% 
    kable_styling(latex_options = "HOLD_position")

```

\newpage

```{r }
loadd(help.table)
  help.table %>%
  mutate(across(where(is.character))) %>%
  mutate(across(where(is.character), linebreak)) %>%
    kbl(booktabs = TRUE,
        escape = FALSE,
        col.names = linebreak(c("Code or category", "Inclusion criteria and definitions", "Examples"), align = "c"),
        caption = "Codes assigned to participant responses during qualitative analysis of the “what helps you to engage in OS?” prompt. Note that many of the examples used could be, and ultimately were, assigned to two or more of the codes."
        ) %>% 
    column_spec(column = 1, width = "50px") %>% 
    column_spec(column = 2, width = "200px") %>% 
    column_spec(column = 3, width = "150px") %>% 
    kable_styling(latex_options = "HOLD_position")

```

\newpage

### Which OS aspects have practitioners encountered in their own formal education?





```{r results-q3-2}
loadd(result_Q3.2)

result_Q3.2 %>% 
  select(-Action) %>% 
  kbl(booktabs = TRUE, 
      digits = 2,
      caption = "Parameter estimates, SE and z-values for covariates in models used to test prediction 3.2") %>% 
  pack_rows("use data", 1, 3) %>% 
  pack_rows("use code", 4, 6) %>%
  pack_rows("use open publication", 7, 8) %>% 
  pack_rows("share data", 9, 11) %>%
  pack_rows("share code", 12, 14) %>% 
  pack_rows("share open publicaiton", 15, 16) %>%
  pack_rows("use edu tools", 17, 19) %>%
  pack_rows("do open review", 20, 21) %>%
  pack_rows("do outreach", 22, 23) %>%
  kable_styling(latex_options = "HOLD_position")
  
```

\newpage


```{r results-q3-3}
loadd(result_Q3.3)

result_Q3.3 %>% 
  kbl(booktabs = TRUE, 
      digits = 2,
      caption = "Parameter estimates, SE and z-values for covariates in models used to test prediction 3.3") %>% 
  pack_rows("global", 1, 9) %>% 
  pack_rows("final", 10, 17) %>%
  kable_styling(latex_options = "HOLD_position")
  
```

\newpage



## FIRST PAGE: CONSENT

### About the study

| This is the first step of a three-part survey of the participants of the Living Norway 2020 Colloquium. The aim of the survey is to investigate perceptions of open science, open science applications, as well as conditions that either promote or limit open science practices. Further, we are interested in understanding how perceptions of and attitudes towards open science develop during the colloquium, as well as the participants’ feedback on the Living Norway Colloquium itself.

| This will be done through three linked surveys sent out before, during, and just after the colloquium.

| To link the three survey parts together at the end of the workshop, we ask you to provide your email address as a temporary identifier for this purpose only. Once the surveys are linked, the email addresses will be immediately and permanently removed from the survey dataset and survey participants will be unidentifiable.

| The University of Bergen (UiB) is responsible for conducting the survey and will ensure that all submissions are handled confidentially. Data will be stored on a safe server and will only be accessible to researchers responsible for summarizing the survey data on behalf of Living Norway. None of the respondents will be identifiable in the final data or in any communication of the results from the study.

| The surveys are important to us in developing Living Norway Colloquia and in better understanding the research community’s views and opinions of open science issues. As this study requires the full consent of participants, you are free to withdraw at any time without the need to justify.

| The study is registered in the UiB data and privacy framework RETTE. 

### Your rights
| As long as you are identifiable in the collected data, you have the right to the following:

- View of your registered personal data.
- Correction of your registered personal data.
- Deletion of your registered personal data.
- Access to a copy of your registered personal data.
- File a complaint for the UiB privacy agent (persornvernombud) or the Norwegian Data Protection Authority (Datatilsynet) regarding the handling of your registered personal data.

| If you have any questions about the study, please contact

- Professor Vigdis Vandvik, bioCEED/UiB, vigdis.vandvik@uib.no
- Christian Bianchi Strømme, bioCEED/UiB, christian.stromme@uib.no

###Consent

**By answering the following survey questions, I confirm that I have received information about the study and that my participation is voluntary.**

**In order to link your answers, we kindly ask you to provide your email address in the field below. Please use the same address on all parts of the survey:**
